---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# mpm

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- [![CRAN status](https://www.r-pkg.org/badges/version/mpm)](https://CRAN.R-project.org/package=mpm) -->
[![R-CMD-check](https://github.com/parunjodhi/mpm/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/parunjodhi/mpm/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

## Purpose 

The goal of this package is to provide a centralized place to clearly visualize the effects of climate change based on key demographics by country. `mpm` is a package that contains data about various greenhouse gas (GHG) emissions by country based on GDP and population.

## Target audience

This package would find excellent use by governmental agencies to determine how their GHG emissions are affecting their communities. 

## Installation

You can install the development version of mpm like so:

``` r
devtools::install_github("parunjodhi/mpm")
```

## Datasets Included
  + `owid_ghg`: A dataset looking at CO2 and Greenhouse Gas (GHG) Emissions filtered for countries in North America.

## Functions Included
  + `population_plot`
  + `gdp_plot`
  + `ghg_plot`
  + `all_demographics`
  
## Basic Usage
These are a few examples of how to use the package and its functions to wrangle and visualize data related to demographics and emissions in countries in North America.

Load `mpm` R package
```{r}
library(mpm)
```

### Data Visualization functions
  1. Visualize the trend in population for Guatemala from 1980 to 2000
```{r}
population_plot("Guatemala", 1980, 2000)
```

  2. Visualize the trend in GDP for Canada from 1940 to 2020
```{r}
gdp_plot("Canada")
```

  3. Visualize the trend in total GHG emissions for the United States from 1940 to 2020
```{r}
ghg_plot()
```

  4. Visualize the trends in population, GDP and total GHG emissions for Jamaica from 2000 to 2020
```{r}
all_demographics("Jamaica", min_year = 2000)
```

## Plan for Phase III

In Phase III, we plan to expand this package to include a function that has the ability to plot how GHG emissions affect health issues (asthma, chronic diseases, air quality) in a given community based on the variables we have currently included.
