---
title: "mpm"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{mpm}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, warning = FALSE, message = FALSE, include = FALSE}
library(dplyr)
library(ggplot2)
```

The `mpm` package contains data on CO2 and Greenhouse Gas (GHG) emissions, pollution (air, household and outdoor pollution) in terms of Disability-Adjusted Life Years (DALYs) and internal displacement within countries.

## The Why?
We wanted to create a centralized package that, in a quick glance, could give users a picture of climate change in any North American country and see how it is affecting the people of that country.

## For Who?
This package would find excellent use by governmental agencies to determine how their GHG emissions are affecting their communities.

## Datasets Included
  - `owid_ghg`: dataset on CO2 and GHG emissions from 1940-2020
  - `air_data`: dataset on different types of air pollution measured in DALYs from 1990-2019
  - `regional_grouping`: dataset grouping North American countries into regions by economic class
  - `displaced_by_disaster`: dataset on internal displacement counts due to extreme events caused by climate change.

The `mpm` datasets only look at North American countries which include: United States, Mexico, Canada, Bermuda, Guatemala, Cuba, Haiti, Dominican Republic, Honduras, Nicaragua, El Salvador, Costa Rica, Panama, Jamaica, Trinidad and Tobago, Belize, Bahamas, Barbados, Saint Lucia, Grenada, Saint Vincent and the Grenadines, Antigua and Barbuda, Dominica, and Saint Kitts and Nevis. 

The `owid_ghg`, `air_data` and `displaced_by_disaster` datasets also include the total values of all the countries for each variable in a row value *North America*.

## How can we use this package?

A number of different climate-change related questions can be answered using the tools this package provides us. Some include:

  1. Compare current and predicted total GHG emissions among different North American countries.
  2. How did the Great Recession of 2008 affect a country?
  3. Given a certain year, would a country have exceeded a given threshold in terms of total GHG, holding everything else constant?
  4. How did a natural disaster in recent years affect internal displacement?

  
### Question 1
**Compare current and predicted total GHG emissions among different North American countries.**

Let's say climate activists in Honduras are arguing that Honduras has faced disproportionate consequences due to natural disasters (arguably exacerbated by climate change) in comparison to the United States who puts out more GHG emissions.

Therefore, let's look at the 2019 level of GHG emitted in both Honduras and the United States.

```{r}
owid_ghg %>%
  filter(country == "Honduras") %>%
  filter(year == 2019)

```

```{r}
owid_ghg %>%
  filter(country == "United States") %>%
  filter(year == 2019)

```

The total GHG emissions in Honduras in 2019 is 28.1 million tonnes. The total GHG emissions in the United States in 2019 is 5771 million tonnes, which is 200 times more that what Honduras emitted for that year. 

The activists can also look at internal displacement pattern in Honduras.

```{r}
displaced_linegraph("Honduras")
```

There is a substantial increase in displaced individuals in Honduras after 2019 that can be attributed to strong hurricanes they face in the coming years.

What happens if no additional actions are taken to curb the GHG emissions? For example, what would then be the level of total GHG emissions for these 2 countries in 2030?

```{r}
ghg_predict(2030, "Honduras")
```

```{r}
ghg_predict(2030)
```

Holding all else constant, in 2030, the predicted GHG emissions in Honduras is estimated to be 32.8 million tonnes while that for the United States is estimated to be 5990 million tonnes.

The climate activists can therefore use some of the tools above to help prove their argument.

### Question 2
**How did the Great Recession of 2008 affect a country?**

Let's look at what happened in and around 2008 in Mexico.

```{r}
all_data_country("Mexico") %>%
  filter(year >= 2000) %>%
  head(20)
```

From the table generated above, we can look at the trends of each variable before, in and after 2008 and make inferences about how  the Great Recessions affected these metrics in Mexico.

### Question 3
**Will a country have exceeded a given threshold in terms of total GHG in a certain year, holding everything else constant?**

Let's say the hard threshold in terms of GHG emissions is 5000 million tonnes (fictional number). Will Canada exceed this threshold in 2040?

```{r}
ghg_predict(2040, "Canada")
```

In 2040, holding all else constant, Canada is estimated to release 1094 million tonnes of GHG emissions. This is still below our fictional threshold level. However, as the warning given in the output states, this prediction is based off of a linear model and not off of an exponential one which may be more likely.

### Question 4
**How did a natural disaster in recent years affect internal displacement?**

If we take a look at the internal displacement of the Cuba following Hurricane Irma in 2017, is there a large spike?

```{r}
displaced_linegraph("Cuba")
```

In 2017, there was a large spike in the total internally displaced individuals. 
